\chapter{Introduction}
\label{chap:introduction}

\section{Motivation}
\label{sec:introduction-motivation}

Epilepsy is a neurological disorder characterized by recurrent seizures caused
by abnormal patterns of brain activity. For many patients, seizures occur
unexpectedly and can interfere with daily activities, safety, independence, and
quality of life. The ability to better understand the brain dynamics that
precede seizures is therefore an important scientific and clinical objective.

Electroencephalography (EEG) provides a non-invasive way to measure brain
activity with high temporal resolution. Because epileptic seizures involve
changes in the temporal organization of neural activity, EEG is a natural
modality for studying seizure-related brain dynamics. In particular, scalp EEG
datasets such as CHB-MIT have become widely used for computational studies of
epileptic seizure analysis and seizure prediction \cite{Tsiouris2018}.

Many computational approaches to seizure analysis focus on extracting signal
features from EEG recordings and using them for classification or prediction.
However, seizure generation is not only a local signal phenomenon. It can also
be understood as a network-level process, where interactions between brain
regions change over time. This motivates the use of functional connectivity and
graph theory, which allow EEG recordings to be represented as networks whose
nodes correspond to recording channels and whose edges represent statistical
relationships between signals \cite{Friston1994,Rubinov2010}.

This thesis investigates whether graph-theoretical features of scalp EEG
functional connectivity change before seizure onset. The work is inspired by
previous graph-theoretical studies of pre-seizure brain network organization,
especially the study by Vecchio et al., which analyzed pre-seizure network
architecture using intracranial EEG recordings \cite{Vecchio2016}. In contrast,
the present thesis applies a related network-analysis framework to scalp EEG
from the CHB-MIT dataset, using a fixed bipolar montage, multiple connectivity
methods, multiple frequency bands, and full-dataset exploratory statistical
analysis.

Prediction remains the broader motivation, but the immediate
goal is to determine whether preictal EEG networks show measurable and
statistically supported changes before seizure onset.

\section{Epileptic Seizures and the Need for Preictal Analysis}
\label{sec:introduction-preictal-analysis}

An epileptic seizure is typically preceded by a period during which brain
activity may begin to deviate from interictal background activity. This period
is often referred to as the preictal period. If measurable preictal changes can
be identified, they may provide useful information about how seizures emerge and
may eventually support seizure-warning or prediction systems.

The distinction between interictal and preictal activity is important in
seizure analysis. The interictal state refers to brain activity between
seizures, away from seizure onset, whereas the preictal state refers to the
period before a seizure during which measurable changes related to seizure
generation may occur. Identifying this transition is one of the central
challenges of seizure prediction, because preictal changes may be subtle,
patient-specific, variable across seizures, and not always visible in the raw
EEG signal \cite{Mormann2007,Kuhlmann2018}. For this reason, computational
features are often used to summarize EEG activity in a more structured way and
to test whether preictal periods differ from interictal baseline activity.

Previous seizure prediction studies using CHB-MIT have shown that the dataset
is useful but also heterogeneous. Recordings differ across subjects, channel
availability is not always consistent, and seizure counts vary substantially
between case folders \cite{Tsiouris2018}. These factors make careful dataset
screening and methodological consistency essential. In this thesis, only
seizures with sufficient preictal duration and the required EEG channels were
included, so that each retained seizure could be analyzed using the same
temporal-window structure and the same graph nodes.

The temporal design of this thesis follows the idea that network changes may be
examined over consecutive time windows before seizure onset. Each seizure was
represented using an interictal Baseline window and three preictal windows:
T0, T1, and T2. This design was motivated by the pre-seizure graph-theoretical
framework of Vecchio et al., who analyzed network organization across
successive three-minute windows before seizure onset \cite{Vecchio2016}. By
using comparable temporal windows, the present thesis allows the scalp EEG
results to be discussed in relation to this earlier intracranial EEG study.

\section{EEG Functional Connectivity and Graph-Theoretical Analysis}
\label{sec:introduction-connectivity-graph-analysis}

Functional connectivity describes statistical relationships between signals
recorded from different brain regions or sensors. In EEG analysis, functional
connectivity can be estimated between pairs of channels and used to construct a
connectivity matrix. Each entry of the matrix represents the estimated coupling
between two EEG signals. This matrix can then be interpreted as a functional
network, where channels are represented as nodes and connectivity values are
represented as edges.

This network representation is useful because it allows seizure-related EEG
activity to be studied not only in terms of individual channels, but also in
terms of the organization of interactions between channels. Graph theory
provides quantitative measures of this organization, including local
segregation, global integration, hub structure, modular organization, and
small-world properties \cite{Rubinov2010}. Such measures have been used in
epilepsy research to study how seizure activity is associated with changes in
brain network topology \cite{Ponten2007,Vecchio2016,Lanzone2022}.

In this thesis, three complementary functional connectivity methods were used:
magnitude-squared coherence, weighted phase-lag index, and amplitude envelope
correlation. Coherence captures frequency-specific linear coupling between EEG
signals. Weighted phase-lag index emphasizes consistent non-zero phase-lag
relationships and is less sensitive to zero-lag coupling that may arise from
volume conduction \cite{Vinck2011}. Amplitude envelope correlation captures
co-fluctuations in band-limited signal amplitudes \cite{Bruns2000,Hipp2012}.
Using multiple connectivity methods makes it possible to examine whether
observed network changes depend on the type of coupling used to define the
network.

The analysis was also performed separately across frequency bands. This is
important because seizure-related network organization may not be identical
across slow and fast EEG rhythms. Previous epilepsy network studies have
reported frequency-specific changes, and Vecchio et al. analyzed pre-seizure
network properties separately across delta, theta, alpha, and beta sub-bands
\cite{Ponten2007,Vecchio2016}. The present thesis therefore uses six frequency
bands: delta, theta, alpha1, alpha2, beta1, and beta2.

After connectivity estimation, each matrix was summarized using graph metrics.
The main analyzed metrics included mean connectivity, clustering coefficient,
global efficiency, characteristic path length, small-worldness, modularity, mean
betweenness centrality, and assortativity. Together, these measures describe
different aspects of functional network organization. For example, global
efficiency measures how efficiently information can be exchanged across the
network, clustering coefficient reflects local connectedness, and modularity
captures the degree to which the network is organized into communities
\cite{Rubinov2010}.

\section{Aim and Research Questions}
\label{sec:introduction-aim-research-questions}

The aim of this thesis is to characterize preictal changes in scalp EEG
functional connectivity networks using graph-theoretical analysis. The study
does not aim to build a final seizure prediction classifier. Instead, it
performs a full-dataset exploratory statistical analysis to determine whether
network metrics differ between interictal baseline activity and preictal
windows before seizure onset.

The main research question is:

\begin{quote}
    Do graph-theoretical features of scalp EEG functional connectivity networks
    change across temporal windows before epileptic seizure onset?
\end{quote}

This question is divided into the following sub-questions:

\begin{enumerate}
    \item Can the CHB-MIT dataset be screened and standardized to support a
    consistent seizure-centered graph-theoretical analysis across subjects and
    seizures?

    \item Do graph metrics differ between an interictal Baseline window and the
    preictal windows T0, T1, and T2?

    \item Are any observed temporal-window effects specific to particular
    frequency bands or connectivity methods?

    \item Are seizure-level findings also visible when the analysis is repeated
    at the subject-average level?

    \item How do the observed scalp EEG findings compare with the pre-seizure
    graph-theoretical framework proposed by Vecchio et al. \cite{Vecchio2016}?
\end{enumerate}

These questions guide the methodological design of the thesis. The dataset
screening and preprocessing stages ensure that the same channel definitions and
temporal windows are used across retained seizures. The connectivity and graph
feature stages convert EEG segments into network-level representations. The
statistical analysis then tests whether graph metrics vary across the temporal
windows before seizure onset.

\section{Thesis Contributions}
\label{sec:introduction-contributions}

The main contributions of this thesis are:

\begin{enumerate}
    \item A full-dataset screening procedure for the CHB-MIT EEG dataset,
    identifying seizures suitable for preictal graph-theoretical analysis.

    \item A standardized preprocessing and segmentation pipeline using a fixed
    \num{22}-channel bipolar montage and four seizure-centered temporal windows:
    Baseline, T0, T1, and T2.

    \item A functional connectivity pipeline that computes coherence, weighted
    phase-lag index, and amplitude envelope correlation across six EEG frequency
    bands.

    \item A graph feature extraction pipeline that converts connectivity
    matrices into weighted functional networks and computes graph-theoretical
    metrics describing network integration, segregation, community structure,
    and centrality.

    \item A full-dataset exploratory statistical analysis of graph metrics
    across temporal windows, including seizure-level mixed-effects models,
    planned Baseline-vs-preictal comparisons, preictal trend tests, and
    subject-average sensitivity analysis.

    \item A comparison of the resulting scalp EEG findings with the
    pre-seizure graph-theoretical framework used by Vecchio et al.
    \cite{Vecchio2016}.
\end{enumerate}

Together, these contributions provide a complete computational workflow for
characterizing preictal network changes in scalp EEG. The resulting analysis is
intended as a foundation for future work on subject-independent seizure
prediction, but the present thesis focuses on exploratory network
characterization and statistical interpretation.

\section{Thesis Structure}
\label{sec:introduction-thesis-structure}

The remainder of this thesis is organized as follows.

\textbf{\Cref{chap:background-state-of-art}} introduces the background and
state of the art. It reviews epilepsy and EEG, seizure prediction and preictal
analysis, functional connectivity, graph-theoretical network analysis, and
previous studies that motivated the present work.

\textbf{\Cref{chap:dataset-preprocessing}} describes the CHB-MIT dataset,
dataset screening, preprocessing, channel selection, temporal window definition,
and frequency-band decomposition.

\textbf{\Cref{chap:connectivity-graph-features}} presents the functional
connectivity and graph feature extraction pipeline. It explains how EEG windows
were transformed into connectivity matrices and how graph-theoretical metrics
were computed.

\textbf{\Cref{chap:statistical-analysis}} describes the statistical methodology
used to test temporal-window effects, Baseline-vs-preictal contrasts, preictal
trends, and subject-average sensitivity analyses.

\textbf{\Cref{chap:experimental-results}} reports the experimental results of
the full-dataset analysis, including the main FDR-corrected findings and the
subject-average sensitivity results.

\textbf{\Cref{chap:discussion}} discusses the interpretation of the findings,
their relationship to previous work, especially Vecchio et al.
\cite{Vecchio2016}, and the methodological limitations of the study.

\textbf{\Cref{chap:conclusion-future-work}} summarizes the thesis and outlines
directions for future work, including the possible extension from exploratory
network characterization to seizure prediction models.