\chapter{Background and State of the Art}
\label{chap:background-state-of-art}

\section{Epilepsy and EEG}
\label{sec:background-epilepsy-eeg}

Epilepsy is a neurological disorder characterized by an enduring predisposition
to generate epileptic seizures. According to the International League Against
Epilepsy, epilepsy is not defined only by the occurrence of a single seizure,
but by a clinical condition in which the probability of further seizures is
sufficiently high or in which an epilepsy syndrome has been diagnosed
\cite{Fisher2014}. An epileptic seizure corresponds to a transient occurrence
of signs or symptoms caused by abnormal excessive or synchronous neuronal
activity in the brain.

Electroencephalography (EEG) is one of the main techniques used to study
epileptic activity. It records voltage fluctuations produced by brain activity
through electrodes placed on the scalp or, in invasive studies, directly inside
or on the surface of the brain. Scalp EEG is non-invasive and provides high
temporal resolution, making it suitable for studying rapid changes in brain
activity around seizure onset. However, scalp EEG is also affected by several
limitations, including volume conduction, movement artifacts, muscle activity,
electrode noise, and variability in electrode montages across recordings.

\section{Seizure Prediction and Preictal State Detection}
\label{sec:background-seizure-prediction}

Seizure prediction aims to identify a warning period before seizure onset. A
successful prediction system would distinguish preictal activity from interictal
activity early enough to allow a useful intervention. This problem has motivated
many studies using EEG features, machine learning, and deep learning methods.
However, seizure prediction remains challenging because preictal patterns may be
weak, patient-specific, variable across seizures, and difficult to separate from
normal fluctuations in interictal EEG activity \cite{Mormann2007,Kuhlmann2018}.

A key difficulty is the definition of the preictal period. There is no
universally accepted preictal duration that applies to all patients and all
seizure types. Some studies define the preictal period as several minutes before
seizure onset, while others use longer intervals. The choice of preictal window
affects feature extraction, statistical testing, and the interpretation of
results. For this reason, studies must clearly define their temporal windows and
avoid treating all pre-seizure activity as a single homogeneous state.

In this thesis, seizure prediction is the broader motivation, but the main
objective is not to build a final classifier. Instead, the focus is on
preictal network characterization. The analysis asks whether graph-theoretical
features of scalp EEG functional connectivity differ between an interictal
Baseline window and consecutive preictal windows. This exploratory step is
important because a prediction model should ideally be motivated by measurable
and interpretable preictal changes.

\section{The CHB-MIT Scalp EEG Dataset}
\label{sec:background-chbmit}

The CHB-MIT scalp EEG dataset is widely used in computational studies of
epileptic seizure analysis. It contains long-term scalp EEG recordings from
pediatric epilepsy cases, together with seizure annotations. Because the data
are publicly available and include multiple subjects and seizures, the dataset
has been used in many seizure detection and seizure prediction studies
\cite{Tsiouris2018}.

Despite its usefulness, the CHB-MIT dataset presents methodological challenges.
The number of seizures differs across case folders, and not all recordings have
the same available channels. Some recordings also differ in montage structure or
contain duplicated channel labels. These issues are important for graph-based
analysis because graph nodes must have consistent meaning across subjects,
seizures, and time windows. If different recordings use different channel sets,
then graph metrics may become difficult to compare.

Previous studies using CHB-MIT have addressed this problem by selecting fixed
subsets of channels that are consistently available across recordings
\cite{Tsiouris2018}. The present thesis follows the same general principle of
montage standardization, but uses a fixed \num{22}-channel bipolar montage after
screening. Recordings and seizures were retained only when the required channels
and sufficient preictal duration were available. This ensured that all retained
connectivity matrices had the same node definitions and could be compared
across the full analysis pipeline.

\section{Functional Connectivity in EEG}
\label{sec:background-functional-connectivity}

Functional connectivity refers to statistical dependence between signals
recorded from different brain regions or sensors \cite{Friston1994}. In EEG
analysis, functional connectivity is often estimated between pairs of channels.
The resulting pairwise values can be organized into a connectivity matrix, where
each matrix entry represents the estimated coupling between two EEG signals.

Different connectivity measures capture different types of relationships.
Coherence measures frequency-specific linear coupling between two signals and
has been widely used in EEG analysis. It is useful when the research question
concerns whether two signals share oscillatory activity in a given frequency
band. However, coherence can be influenced by volume conduction and common
reference effects, because zero-lag correlations may arise from signal mixing
rather than true physiological interaction.

Phase-based measures were developed partly to reduce this problem. The weighted
phase-lag index (wPLI) emphasizes consistent non-zero phase-lag relationships
and is less sensitive to zero-phase coupling caused by volume conduction
\cite{Vinck2011}. This makes wPLI useful for EEG connectivity analysis, where
spurious zero-lag relationships are a known concern. Nevertheless, wPLI captures
only a specific aspect of coupling and may not reflect all forms of functional
interaction.

Amplitude-based methods provide another complementary view. Amplitude envelope
correlation (AEC) measures the correlation between the envelopes of
band-limited signals and can capture co-fluctuations in oscillatory power
\cite{Bruns2000,Hipp2012}. This can reveal relationships that are not
necessarily phase-locked. In the present thesis, coherence, wPLI, and AEC were
used together to examine whether preictal network changes depend on the
connectivity definition.

Connectivity analysis is also strongly frequency-dependent. EEG rhythms are
commonly analyzed in bands such as delta, theta, alpha, and beta. Seizure-related
changes may appear in some frequency ranges but not others. For this reason, the
present thesis computes connectivity separately in six frequency bands: delta,
theta, alpha1, alpha2, beta1, and beta2. This follows the frequency-band
structure used in the pre-seizure graph-theoretical study by Vecchio et al.
\cite{Vecchio2016}.

\section{Graph-Theoretical Analysis of Brain Networks}
\label{sec:background-graph-theory}

Graph theory provides a mathematical framework for studying networks. In a
functional EEG network, nodes represent EEG channels and edges represent
functional connectivity values between channel pairs. Once a connectivity matrix
has been interpreted as a graph, graph-theoretical measures can be used to
summarize different aspects of network organization \cite{Rubinov2010}.

Several graph metrics are commonly used in brain network analysis. Clustering
coefficient describes the tendency of neighboring nodes to be connected to one
another and is often interpreted as a measure of local segregation. Global
efficiency measures how efficiently information can be exchanged across the
network and is related to network integration \cite{Latora2001,Rubinov2010}.
Characteristic path length describes the average shortest-path distance between
nodes. Modularity measures the degree to which a network can be divided into
communities or modules. Betweenness centrality identifies nodes that lie on many
shortest paths and may therefore act as bridges within the network.

Small-world organization is another important concept in brain network
analysis. Small-world networks combine high local clustering with relatively
short path lengths, providing a balance between local specialization and global
integration \cite{Watts1998}. Brain networks have often been studied using this
framework because efficient neural communication may require both local and
distributed organization \cite{Rubinov2010}. In epilepsy, changes in
small-worldness, efficiency, or centrality may reflect altered synchronization
or reorganization of functional brain networks.

When applying graph theory to EEG connectivity, methodological choices strongly
affect the results. These include the connectivity measure, frequency band,
thresholding strategy, treatment of negative values, and handling of disconnected
graphs. Therefore, graph-theoretical results must be interpreted in relation to
the full pipeline that produced them. In this thesis, the same preprocessing,
connectivity estimation, thresholding, and graph metric procedures were applied
to all retained seizures to support consistent comparison across temporal
windows.

\section{Graph-Theoretical Studies in Epilepsy}
\label{sec:background-graph-epilepsy}

Graph-theoretical approaches have been used to study epilepsy because seizures
can be understood as network phenomena involving abnormal synchronization and
altered communication between brain regions. Rather than focusing only on signal
amplitude or spectral power at individual electrodes, graph analysis makes it
possible to examine how functional interactions are organized across the
recording montage.

Ponten et al. used graph-theoretical analysis to study intracranial EEG
recordings during mesial temporal lobe seizures and reported seizure-related
changes in network measures such as clustering and path length
\cite{Ponten2007}. This work helped establish the relevance of graph theory for
studying epileptic brain dynamics. Later studies extended graph analysis to
different EEG settings, connectivity measures, and clinical questions.

Vecchio et al. performed one of the main studies motivating the present thesis.
They analyzed pre-seizure network architecture using intracranial EEG and
computed graph-theoretical measures across consecutive three-minute windows
before seizure onset \cite{Vecchio2016}. Their study examined whether the local
network organization of the epileptic focus changed before seizures and reported
frequency-specific differences in small-world-related properties. The present
thesis adopts a related temporal-window idea but applies it to scalp EEG from
CHB-MIT rather than intracranial EEG.

More recent work has also applied graph metrics to scalp EEG in epilepsy.
Lanzone et al. used EEG connectivity and graph-theoretical measures to study
changes associated with vagus nerve stimulation cycles in drug-resistant
epilepsy \cite{Lanzone2022}. Their work illustrates how graph metrics such as
small-worldness, betweenness centrality, and global efficiency can be used to
summarize functional EEG network organization in clinical epilepsy research.

Together, these studies support the idea that epilepsy can be investigated as a
network disorder and that graph-theoretical metrics may reveal changes not
captured by single-channel EEG features alone. However, results depend on the
recording modality, patient population, connectivity method, frequency band,
and statistical design. This creates a need for carefully documented pipelines
when graph theory is applied to public scalp EEG datasets.

\section{Positioning of the Present Thesis}
\label{sec:background-positioning}

The present thesis is positioned at the intersection of scalp EEG analysis,
functional connectivity, graph theory, and exploratory preictal seizure
characterization. It is inspired by graph-theoretical studies of pre-seizure
network organization, especially Vecchio et al. \cite{Vecchio2016}, but differs
from them in several important ways.

First, the present study uses scalp EEG from the CHB-MIT dataset rather than
intracranial EEG. Scalp EEG is non-invasive and more widely available, but it
has lower spatial specificity and is more affected by volume conduction and
artifacts. This makes it important to use multiple connectivity measures and to
interpret graph metrics cautiously.

Second, the analysis is performed across the screened full dataset rather than
on a single subject or a small manually selected subset. This required a
systematic screening procedure to ensure that retained seizures had sufficient
preictal duration and the required EEG channels. The analysis therefore
preserves the relationship between subjects, seizures, temporal windows,
frequency bands, connectivity methods, and graph metrics.

Third, the thesis does not present a final seizure prediction classifier.
Instead, it asks whether preictal network changes can be detected in a
statistically interpretable way. This distinction is important because
classification performance alone does not necessarily explain which aspects of
EEG network organization change before seizures. By focusing on graph metrics
and statistical testing, the thesis provides an exploratory basis for future
prediction work.

The methodological contribution of the thesis is therefore a complete and
reproducible pipeline for transforming CHB-MIT scalp EEG recordings into
functional connectivity networks, extracting graph-theoretical metrics, and
testing temporal-window effects before seizure onset. The scientific objective
is to evaluate whether these network measures reveal consistent preictal
alterations and how such findings compare with previous graph-theoretical
epilepsy studies.