\chapter{Discussion}
\label{chap:discussion}

\section{Overview of the Main Finding}
\label{sec:discussion-main-finding}

This thesis investigated whether graph-theoretical features of scalp EEG
functional connectivity change across temporal windows before epileptic seizure
onset. The analysis was exploratory and descriptive in nature: it was designed
to characterize preictal network organization, not to establish a clinical
biomarker or to build a seizure prediction system.

After correction for multiple comparisons, the main finding was a significant
temporal-window effect for global efficiency computed from delta-band coherence
networks. This effect survived FDR correction in both the seizure-level omnibus
analysis and the subject-average omnibus analysis. At the seizure level, the
planned post-hoc comparisons localized the strongest corrected difference to
the Baseline-vs-T1 contrast, where global efficiency was higher in T1 than in
the Baseline window. No preictal T0--T1--T2 trend survived FDR correction.

This result suggests that, within the present pipeline, the most consistent
preictal alteration was not a broad change across all graph measures,
connectivity methods, or frequency bands. Instead, the corrected evidence was
specific to one network feature: global efficiency of delta-band coherence
networks. Therefore, the result should be interpreted as a focused exploratory
finding rather than as evidence for a general preictal network signature.

\section{Interpretation of Delta-Band Coherence Global Efficiency}
\label{sec:discussion-delta-coherence-efficiency}

Global efficiency is commonly interpreted as a measure of network integration,
because it reflects how efficiently information can be exchanged across the
network through short paths \cite{Latora2001,Rubinov2010}. In the present
analysis, higher global efficiency in delta-band coherence networks during T1
indicates that, after graph construction and thresholding, the retained
functional network had a more globally integrated structure than during the
interictal Baseline window.

This finding is compatible with the general idea that seizure generation may be
associated with changes in large-scale functional organization before clinical
onset. However, the result should be interpreted cautiously. The networks in
this thesis are sensor-level scalp EEG networks, where nodes correspond to
bipolar EEG channels rather than directly to anatomical brain regions. Moreover,
coherence may be influenced by volume conduction and common signal components.
For this reason, the result is best described as a change in scalp EEG
functional network organization, not as direct evidence of a specific anatomical
mechanism.

The frequency specificity of the result is also relevant. The corrected effect
appeared in the delta band, while no other frequency band survived correction
for the same metric and method. This suggests that slow-frequency coherence may
carry information about preictal network changes in this dataset. Previous
epilepsy network studies have also reported frequency-dependent network changes,
although the specific bands and metrics differ across studies
\cite{Ponten2007,Vecchio2016,Lanzone2022}. Therefore, the present result fits
the broader idea that seizure-related network organization may depend on
frequency band, but it should not be generalized beyond the tested dataset and
pipeline.

\section{Comparison with Vecchio et al.}
\label{sec:discussion-comparison-vecchio}

The present thesis was inspired by the pre-seizure graph-theoretical study of
Vecchio et al. \cite{Vecchio2016}. Their work analyzed the architecture of
local connections around the epileptic focus using intracranial EEG. They
divided the pre-seizure period into consecutive three-minute windows and
computed coherence-based graph measures across several frequency bands. This
temporal structure directly motivated the Baseline, T0, T1, and T2 design used
in the present thesis.

There are important similarities between the two studies. Both approaches treat
seizure development as a network-level process rather than only as a
single-channel signal phenomenon. Both use coherence as one of the central
connectivity measures, analyze graph-theoretical features across frequency
bands, and examine whether network organization changes before seizure onset.
In this sense, the present thesis extends the same general idea to a different
recording setting: scalp EEG from the CHB-MIT dataset.

However, the methodological differences are substantial. Vecchio et al. used
intracranial stereo-EEG recordings and focused on local contacts near the
epileptic focus \cite{Vecchio2016}. In contrast, the present thesis used scalp
EEG and a fixed \num{22}-channel bipolar montage. Therefore, the networks in
the two studies represent different physiological and spatial levels.
Intracranial EEG provides more local and spatially specific information, while
scalp EEG is non-invasive but more affected by volume conduction, artifacts, and
sensor-level mixing.

The main findings also differ. Vecchio et al. reported pre-seizure changes in
small-world-related network organization, especially in theta and alpha2
frequency ranges \cite{Vecchio2016}. In the present thesis, the strongest
FDR-corrected effect was not small-worldness but global efficiency, and it was
observed in delta-band coherence networks. This difference may reflect the
recording modality, the use of scalp rather than intracranial EEG, the larger
and more heterogeneous dataset, the fixed whole-scalp sensor montage, the graph
construction choices, or the statistical correction across many
metric--method--band combinations.

Therefore, the present results should not be viewed as a direct replication of
Vecchio et al. Instead, they provide a related scalp EEG analysis inspired by
their temporal and graph-theoretical framework. Both studies support the
broader idea that network organization can change before seizure onset, but
they point to different specific graph features and frequency bands. This
underlines the importance of recording modality, preprocessing, connectivity
definition, graph construction, and statistical design when comparing
graph-theoretical epilepsy studies.

\section{Seizure-Level and Subject-Average Results}
\label{sec:discussion-seizure-subject-average}

The primary seizure-level analysis used seizures as repeated-measures units
while accounting for the fact that multiple seizures could come from the same
subject/case identifier. This analysis found a significant omnibus window
effect for delta-band coherence global efficiency and localized the strongest
post-hoc effect to Baseline vs T1.

The subject-average sensitivity analysis asked whether the same pattern remained
visible when each subject/case identifier contributed one average trajectory.
This is important because seizure counts were not balanced across subjects, and
subjects with many seizures could otherwise have a stronger influence on the
seizure-level analysis. The same metric--method--band combination survived FDR
correction in the subject-average omnibus analysis, supporting the robustness of
the main temporal-window effect at the subject/case level.

However, the subject-average post-hoc comparisons did not survive FDR
correction. This means that the subject-average analysis supported the presence
of an overall temporal-window effect, but did not provide corrected evidence for
one specific Baseline-vs-preictal contrast. A cautious interpretation is that
the Baseline-to-T1 increase was most clearly expressed at the seizure level,
while the subject-average analysis confirmed the general direction of the
effect but with reduced power and increased variability across subject/case
identifiers.

\section{Why Most Metrics Did Not Survive Correction}
\label{sec:discussion-nonsignificant-results}

Most graph metrics, connectivity methods, and frequency bands did not survive
FDR correction. This is an important part of the result rather than a failure of
the analysis. The pipeline tested many combinations of graph metric,
connectivity method, and frequency band. Multiple-comparison correction was
therefore necessary to reduce the risk of false-positive interpretation. Under
this correction, only the strongest and most consistent effect remained
significant.

There are several possible reasons why most effects were not significant. First,
preictal changes may be heterogeneous across subjects and seizures. A graph
metric that changes before some seizures may not change consistently across the
full dataset. Second, scalp EEG may be less sensitive to local epileptogenic
network changes than intracranial EEG. Third, different connectivity measures
capture different aspects of signal dependence, and not all of them may be
sensitive to the same preictal dynamics. Fourth, graph metrics are affected by
thresholding and network construction choices, so weak effects may not be
stable across the full pipeline \cite{VanWijk2010,Garrison2015,Adamovich2022}.

The absence of widespread corrected effects also supports a cautious conclusion:
the present analysis does not show a broad, universal preictal graph-theoretical
signature in CHB-MIT scalp EEG. Instead, it identifies one corrected
metric--method--band combination that merits further investigation.

\section{Methodological Strengths}
\label{sec:discussion-strengths}

A strength of this thesis is that the analysis was applied to the screened
full CHB-MIT dataset rather than only to a single subject or one selected
recording. The screening procedure retained seizures only when sufficient
preictal duration and the required EEG channels were available. This allowed the
analysis to use a consistent temporal-window structure and a fixed node
definition across retained seizures.

Another strength is the use of multiple connectivity methods. Coherence, wPLI,
and AEC capture different types of functional relationships. Coherence captures
frequency-specific linear coupling, wPLI emphasizes non-zero phase-lag
relationships and is less sensitive to zero-lag effects, and AEC captures
co-fluctuations in amplitude envelopes \cite{Vinck2011,Bruns2000,Hipp2012}.
Using multiple methods made it possible to test whether preictal effects were
method-specific rather than assuming that one connectivity definition was
sufficient.

The statistical design is also a strength. The primary analysis preserved the
seizure-level repeated-measures structure and accounted for nesting within
subject/case identifiers. The subject-average sensitivity analysis then tested
whether the main finding remained visible when each subject/case contributed
equally. Finally, FDR correction was applied to reduce the risk of
over-interpreting isolated uncorrected effects.

\section{Limitations}
\label{sec:discussion-limitations}

The study also has several limitations. First, the analysis used scalp EEG,
which has lower spatial specificity than intracranial EEG. Functional networks
constructed from scalp EEG channels should therefore be interpreted as
sensor-level networks, not direct maps of anatomical brain connectivity.

Second, no manual artifact rejection or ICA-based artifact correction was
applied. This decision made the pipeline more reproducible and avoided
subjective manual decisions, but it also means that ocular, muscular, movement,
or electrode artifacts may have influenced some connectivity estimates. This is
especially relevant for scalp EEG connectivity analysis.

Third, coherence-based connectivity can be affected by volume conduction and
common signal components. The inclusion of wPLI partly addressed this concern,
because wPLI is designed to reduce sensitivity to zero-lag coupling
\cite{Vinck2011}. However, the main corrected result was found for coherence,
so volume conduction remains an important limitation when interpreting the
finding.

Fourth, graph metrics depend on methodological choices such as proportional
thresholding, treatment of negative values, and handling of disconnected graphs.
Although the same choices were applied consistently across all matrices, there
is no universally optimal threshold for EEG functional connectivity networks
\cite{VanWijk2010,Garrison2015,Adamovich2022}. The results should therefore be
understood as conditional on the graph construction pipeline used in this
thesis.

Fifth, the CHB-MIT dataset is heterogeneous. Subjects differ in seizure count,
recording duration, channel availability, and clinical characteristics.
Moreover, the CHB-MIT case identifiers do not perfectly correspond to
independent biological subjects, because one case folder is a later recording
from the same individual as another case folder. For this reason, the thesis
uses the term subject/case identifier and interprets group-level results
cautiously.

Finally, this thesis did not build or validate a seizure prediction classifier.
The analysis identifies an exploratory network-level finding, but it does not
demonstrate predictive performance on unseen subjects or seizures. A prediction
model would require additional design choices, including subject-wise train-test
splitting, leakage control, feature selection, model validation, and evaluation
of sensitivity and false-alarm rates \cite{Mormann2007,Kuhlmann2018}.

\section{Implications for Future Work}
\label{sec:discussion-future-work}

The main implication of this thesis is that graph-theoretical analysis of scalp
EEG functional connectivity can reveal statistically supported preictal network
differences in the CHB-MIT dataset, but the evidence is specific rather than
broad. The strongest corrected effect involved delta-band coherence global
efficiency, suggesting that global integration of slow-frequency coherence
networks may be a useful feature to examine in future work.

Future studies could extend this work in several directions. First, artifact
rejection and robustness analyses could be added to test whether the main
finding remains stable after removing noisy epochs or recordings. Second,
threshold sensitivity analyses could examine whether the result is preserved
across different graph densities. Third, alternative graph construction methods
could be tested, including weighted metrics without hard thresholding. Fourth,
the analysis could be repeated with subject-specific models to evaluate whether
preictal graph changes are more consistent within individual patients than
across the full dataset.

A prediction-oriented extension would also be possible. Such work should use a
strict subject-wise validation strategy to avoid data leakage and to test
generalization to unseen subjects. The present findings could guide feature
selection, but they should not be treated as sufficient evidence for clinical
prediction. Instead, they provide an exploratory and interpretable foundation
for future seizure-prediction studies based on functional connectivity and
graph-theoretical EEG features.