% !TEX program = xelatex
\documentclass[a4paper,singleside,12pt]{report}
% \documentclass[a4paper,twoside,12pt]{report} % For printed double-sided version

% University of Bologna thesis style.
% Compile with XeLaTeX because ai_bo_thesis uses fontspec.
\usepackage{ai_bo_thesis}

\usepackage[english]{babel}

% Main font
\setmainfont{Times New Roman}

% Mathematics and symbols
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}

% Tables
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{longtable}

% Figures
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, decorations.pathreplacing, calc}
% Units and numbers
\usepackage{siunitx}

% Text and URLs
\usepackage{xurl}
\usepackage{csquotes}

% Algorithms / pseudocode, if needed later
% \usepackage{algorithm}
% \usepackage{algpseudocode}

% Bibliography
\usepackage[
    backend=biber,
    style=trad-plain,
    sorting=none,
    giveninits=true
]{biblatex}
\addbibresource{bibliography/references.bib}

% Figure path
\graphicspath{{figures/}}

% Hyperref should usually be loaded near the end of the preamble
\usepackage[hidelinks]{hyperref}

% Clever references after hyperref, optional but useful
\usepackage[nameinlink,noabbrev]{cleveref}

\begin{document}

    % ----------------------------------------------------------------------
    % Frontispiece metadata
    % ----------------------------------------------------------------------

    \title{Graph-Theoretic Analysis of Scalp EEG Functional Connectivity for Epileptic Seizure Prediction}

    \topic{Cognition and Neuroscience}

    \candidate{Leila Sadakbayeva}

    % Confirm final supervisor / co-supervisor roles before submission.
    \supervisor{Prof.~Francesca Starita}
    \cosupervisor{& Prof.~Saverio Giallorenzo}

    \academicyear{2025-2026}

    % Update according to the official graduation session.
    \session{to be confirmed}

    \frontispiece
    \newpage
    \begin{flushright}
    \emph{
    \LARGE{To my brother,}\\\vspace{2mm} 
    \LARGE{whose strength and courage}\\\vspace{2mm} 
    \LARGE{in living with epilepsy}\\\vspace{2mm} 
    \LARGE{inspired this work.}\\\vspace{2mm} 
    }
    \end{flushright}

    % ----------------------------------------------------------------------
    % Abstract
    % ----------------------------------------------------------------------
    
    \chapter*{Abstract}
    Epileptic seizures may be preceded by subtle changes in brain activity that are
    not always visible in raw EEG signals. This thesis investigates whether such
    preictal changes can be characterized using graph-theoretical features of scalp
    EEG functional connectivity in the CHB-MIT dataset.

    A complete analysis pipeline was developed to screen the dataset, preprocess EEG
    recordings, extract seizure-centered temporal windows, compute functional
    connectivity, construct graph representations, and perform exploratory
    statistical analysis. Each usable seizure was represented by an interictal
    Baseline window and three consecutive preictal windows, T0, T1, and T2.
    Connectivity was estimated across six frequency bands using coherence, weighted
    phase-lag index, and amplitude envelope correlation. The resulting connectivity
    matrices were converted into weighted functional networks and summarized using
    graph metrics.

    Overall, the results suggest that preictal network changes in this dataset were
    most consistently reflected in altered global integration of delta-band
    coherence networks. The thesis provides an exploratory graph-theoretical
    characterization of preictal scalp EEG connectivity and a basis for future
    prediction-oriented work.

    \cleardoublepage

    % ----------------------------------------------------------------------
    % Tables of contents
    % ----------------------------------------------------------------------

    \toc
    \figstoc
    \tablestoc

    \begintext

    % ----------------------------------------------------------------------
    % Main chapters
    % ----------------------------------------------------------------------

    \input{chapters/01_introduction}

    \input{chapters/02_background_state_of_art}

    \input{chapters/03_dataset_preprocessing}

    \input{chapters/04_connectivity_graph_features}

    \input{chapters/05_statistical_anlysis}

    \input{chapters/06_experimental_results}

    \input{chapters/07_discussion}

    % \input{chapters/08_conclusions_future_work}

    % ----------------------------------------------------------------------
    % Appendices
    % ----------------------------------------------------------------------

    \appendix

    \input{appendices/appendix_reproducibility}

    % ----------------------------------------------------------------------
    % Bibliography
    % ----------------------------------------------------------------------

    \printbibliography[heading=bibintoc]

    % ----------------------------------------------------------------------
    % Acknowledgements
    % ----------------------------------------------------------------------

    \acknowledgements

    I would like to thank my supervisors for their guidance and support
    throughout this thesis. I am also grateful to my family, friends, and
    colleagues for their encouragement during the development of this work.

\end{document}